---
title: "p8105_hw1_rs4025"
author: "Rebecca Silva"
date: "9/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1

```{r problem 1, echo = FALSE}
#load packages
library(tidyverse)

#create data frame
df1 = tibble(
  norm_samp = rnorm(8),
  norm_samp_log = norm_samp > 0,
  char_var = as.character(norm_samp), 
  factor_var =  factor( c( "grp1", "grp2", "grp3", "grp3", "grp2", "grp1", "grp2", "grp3" ))
)

#try taking mean of each var
mean(pull (df1, norm_samp))
mean(pull (df1, norm_samp_log))
mean(pull (df1, char_var))
mean(pull (df1, factor_var))

#convert vars 
as.numeric(pull (df1, norm_samp_log))
as.numeric(pull (df1, char_var))
as.numeric(pull (df1, factor_var))
```

When trying to take the mean of each variable before converting, the output for the character and factor vector are 'NA's, and I recieve a warning message. R is able to take the mean of the logical vector because TRUE and FALSE are coded as 1's and 0's, respectively.  

Using 'as.numeric' converts
 * the logicals into integers of 0 (false) and 1 (true).
 * the characters into a numeric value.
 * the factors into the numeric value of their assigned level.

Essentially, the function tries to convert each element of a vector into a number. Without converting the variables into a number, R is unable to take the mean, and therefore we get an error. 

```{r}
#logical to numeric
logical_num = as.numeric (pull (df1, norm_samp_log))

pull( df1, norm_samp) * logical_num

#logical to factor
logical_factor = as.factor (pull (df1, norm_samp_log))

pull( df1, norm_samp) * logical_factor #not possible

#logical to factor to numeric
logical_factor_num = as.numeric( 
  as.factor( pull (df1, norm_samp_log))
)

pull( df1, norm_samp) * logical_factor_num

```

# Problem 2

```{r}
# create dataframe
df2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  log_var = x + y > 1, 
  num_var = as.numeric(log_var),
  factor_var = as.factor(log_var)
)
```

The data frame, 'df2', has `r nrow(df2)` rows and `r ncol(df2)` columns. The mean of x is `r mean( pull(df2, x))`, the median of x is `r median( pull(df2, x))`, and the standard deviation of x is `r sd( pull(df2, x))`. The proportion of cases for which $x+y > 1$ is `r sum( pull(df2, log_var))/500`. 

```{r}
# scatter plot - color by logical 
ggplot(df2, aes(x = x, y = y, color = log_var)) + geom_point()

# scatter plot - color by numeric 
ggplot(df2, aes(x = x, y = y, color = num_var)) + geom_point()

# scatter plot - color by factor 
ggplot(df2, aes(x = x, y = y, color = factor_var)) + geom_point()

#export first scatter plot
ggsave("scatter_plot_hw1.pdf", height = 4, width = 6) 
# ^ make sure goes to project directory 
```

Comment on color scales:


