---
title: "p8105_hw1_rs4025"
author: "Rebecca Silva"
date: "9/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(tidyverse)
```

# Problem 1

```{r df, message = FALSE}
#create data frame
df1 = tibble(
  norm_samp = rnorm(8),
  norm_samp_log = norm_samp > 0,
  char_var = as.character(norm_samp), 
  factor_var =  factor( c( "grp1", "grp2", "grp3", "grp3", "grp2", "grp1", "grp2", "grp3" ))
)
```


```{r, results = 'hide'}
# Q: should we show results)
#try taking mean of each var
mean(pull (df1, norm_samp))
mean(pull (df1, norm_samp_log))
mean(pull (df1, char_var))
mean(pull (df1, factor_var))
```

When trying to take the mean of each variable before converting, the output for the character and factor vector are NA's, and I receive a warning message. We can obtain a mean for the numeric vector, as it is numeric, and the logical vector because TRUE and FALSE are converted to values 1 and 0, respectively, when used in a function.  

```{r, results = 'hide'}
#convert vars 
as.numeric(pull (df1, norm_samp_log))
as.numeric(pull (df1, char_var))
as.numeric(pull (df1, factor_var))
```

Using 'as.numeric' converts

 * the logicals into integers of 0 (false) and 1 (true).
 * the characters into a numeric value.
 * the factors into the numeric value of their assigned level.

Essentially, the function tries to convert each element of the vector into a number. Without converting the variables into a number, R is unable to take the mean, and therefore we get an error. 

```{r}
#logical to numeric
logical_num = as.numeric(pull (df1, norm_samp_log))
pull( df1, norm_samp) * logical_num

#logical to factor
logical_factor = as.factor(pull (df1, norm_samp_log))
pull( df1, norm_samp) * logical_factor #not possible

#logical to factor to numeric
logical_factor_num = as.numeric( 
  as.factor( pull (df1, norm_samp_log)))
pull( df1, norm_samp) * logical_factor_num
```

Similar to when we were trying to find the mean of a factor variable, multiplying a numeric variable by a factor variable will no work since multiplication does not work on anything other than a number. 

# Problem 2

```{r}
# create dataframe
df2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  log_var = x + y > 1, 
  num_var = as.numeric(log_var),
  factor_var = as.factor(log_var)
)
```

The data frame, 'df2', has `r nrow(df2)` rows and `r ncol(df2)` columns. The mean of x is `r mean( pull(df2, x))`, the median of x is `r median( pull(df2, x))`, and the standard deviation of x is `r sd( pull(df2, x))`. The proportion of cases for which $x+y > 1$ is `r sum( pull(df2, log_var))/500`. 

```{r}
# scatter plot 1 - color by logical 
ggplot(df2, aes(x = x, y = y, color = log_var)) + geom_point()

#export first scatter plot
ggsave("scatter_plot_hw1.pdf", height = 4, width = 6) 

# scatter plot 2 - color by numeric 
ggplot(df2, aes(x = x, y = y, color = num_var)) + geom_point()

# scatter plot 3 - color by factor 
ggplot(df2, aes(x = x, y = y, color = factor_var)) + geom_point()
```

The color scales for the factor and logical variables are discrete, using two colors to represent 'False' and 'True' with red and blue, respectively. The color scale for the numeric vector is continuous; darker hues of blue indicate values closer to 0 and lighter hues indicate values of the numeric vector that are closer to 1. 




